Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
more fixes

\newcommand{\Sof}[1]{\textbf{[Sofia:[}{\color{red} #1}\textbf{]]}}
\newcommand{\Luis}[1]{\textbf{[Luis:[}{\color{green} #1}\textbf{]]}}
\newcommand{\Rui}[1]{\textbf{[Rui:[}{\color{blue} #1}\textbf{]]}}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.55\textwidth]{figures/language.pdf}
 	\caption{Maintainability difference by language after security refactorings}
	\label{fig:lang_main}
\end{figure}

In this study, we also evaluated the impact of each programming language on the
maintainability of open-source software, Figure~\ref{fig:lang_main}. Although the
dataset contains security refactorings for more than $13$ languages, we only
present the results for \emph{Java}, \emph{C}, \emph{Ruby} and \emph{PHP} because
one of the requirements of the Wilcoxon signed-rank test is that the input sample
has to have more than $20$ measurements. The other languages were integrated in the
\emph{Others} set. The results of the Wilcoxon signed-rank test by pattern returned
statistical evidence which supports that the maintainability of open-source software
decreases for security refactorings in 3 different languages: \emph{C}
(p = $5.72$x$10^{-9}$), Ruby (p = 0.057) and PHP (p = $3.54$x$10^{-6}$). The results
also provide statistical significance for when security refactorings are performed
in \emph{Java} (p = $3.61$x$10^{-2}$). \emph{Python} and the other languages
that integrate the \emph{Others} set contain more cases where the maintainability
increases. However, the results obtained for these patterns are not statistically
significant.

@@

Projects with \emph{Broken Authentication} flaws are the most affected ones by security
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any
change. The results also show that fixing \emph{Memory Leak}s affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow \emph{Denial-of-Dervice} attacks (e.g., CVE-2014-3506) has a negative impact
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of
the cases exhibit maintainability improvement while 22.5\% (9) show that overall
maintainability does not suffer any change. \emph{Miscellaneous} refactorings pattern
integrates several types of security flaws whose impact is of 43.1\% (84) negative
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not
exhibit any change in maintainability and 25.1\% (49) of cases have a positive
impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting} vulnerabilities and by \emph{Using
Components with Know Vulnerabilities}. \emph{Cross-Site Scripting} does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From 21 cases evaluated for \emph{Using Componentes with Known
Vulnerabilities}, only 4.8\% (1) has a positive effect in the maintainability
of software. 14.3\% (3) harms maintainability software and 81\% has no effect
in the maintainability.

@@

\textbf{\textit{Broken Authentication \& Session Management, Denial-of-Service attacks and Memory Leaks and Miscellaneous patterns are more likely to affect open-source software maintainability.}} Although there are
a few refactorings that harm software maintainability in the \emph{Cross-Site Scripting} and \emph{Using Components with Know Vulnerabilities}, overall maintainability does not suffer any changes for these patterns. For the remaining patterns, \emph{Injection} and \emph{Cross-Site Request Forgery}, results are not statistically significant.

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
\emph{Broken Authentication} and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.

>>>>

%
% Add comments in the text
%
\newboolean{showcomments}
\setboolean{showcomments}{false}
%\setboolean{showcomments}{false}

\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[3]{
  {\color{#2}\small\fbox{\bfseries\sffamily\scriptsize#1}}
  {\color{#2}\sffamily\small$\triangleright~$\textit{\small #3}$~\triangleleft$}
  }
  }
  {\newcommand{\nb}[3]{}
  }

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{green}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.56\textwidth]{figures/language.pdf}
 	\caption{Maintainability difference by language after security refactorings}
	\label{fig:lang_main}
\end{figure}

In this study, we also evaluated the impact of each programming language on the
maintainability of open-source software, Figure~\ref{fig:lang_main}. Although the
dataset contains security refactorings for more than $13$ languages, we only
present the results for \emph{Java}, \emph{C}, \emph{Ruby} and \emph{PHP} because
one of the requirements of the Wilcoxon signed-rank test is that the input sample
has to have more than $20$ measurements. The other languages were integrated in the
\emph{Others} set. The results of the Wilcoxon signed-rank test by pattern returned
statistical evidence which supports that the maintainability of open-source software
decreases for security refactorings in 3 different languages: \emph{C}
(p = $5.72$x$10^{-9}$), Ruby (p = 0.057) and PHP (p = $3.54$x$10^{-6}$). The results
also provide statistical significance for when security refactorings are performed
in \emph{Java} (p = $3.61$x$10^{-2}$). \emph{Python} and the languages
that integrate the \emph{Others} set contain more cases where the maintainability
increases. However, the results obtained for these patterns are not statistically
significant.

@@

Software with \emph{Broken Authentication} flaws are the most affected ones by security
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) yielded a positive impact and only 13.8\% (6) did not suffer any
change. The results also show that fixing \emph{Memory Leak}s affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow \emph{Denial-of-Dervice} attacks (e.g., CVE-2014-3506) has a negative impact
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of
the cases exhibit maintainability improvement while 22.5\% (9) show that overall
maintainability does not suffer any change. \emph{Miscellaneous} refactorings pattern
integrates several types of security flaws whose impact is of 43.1\% (84) negative
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not
exhibit any change in maintainability and 25.1\% (49) of cases have a positive
impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting} vulnerabilities and by \emph{Using
Components with Known Vulnerabilities}. \emph{Cross-Site Scripting} does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From the 21 cases evaluated for \emph{Using Componentes with Known
Vulnerabilities}, only 4.8\% (1) has a positive effect in the maintainability
of software. 14.3\% (3) harms maintainability software and 81\% has no effect
in the maintainability.

@@

\textbf{\textit{Broken Authentication \& Session Management, Denial-of-Service attacks and Memory Leaks and Miscellaneous patterns are more likely to affect open-source software maintainability.}} Although there are
a few refactorings that harm software maintainability in the \emph{Cross-Site Scripting} and \emph{Using Components with Known Vulnerabilities}, overall maintainability does not suffer any changes for these patterns. For the remaining patterns, \emph{Injection} and \emph{Cross-Site Request Forgery}, results are not statistically significant.

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Known Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
\emph{Broken Authentication} and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.
